\sout{Machine Learning for Global Conquest} \\ Multimatrices and Their Derivatives

### 2021-11-04 Math/book.tex
Backing up changes. No idea what's changed.

@@

>>>>

(Side note. I found an alternative definition of orthoganal in this video: \url{https://youtu.be/ACZC_XEyg9U?t=1427}
That said $R^TR = I$ says $R$ is orthoganal because the transformation this matrix represents preserves the inner product
between two vectors. IE, $(Ra)^T(Rb) = a^Tb$. I wonder how this relates.)

@@

The attention mechanisms from \href{https://www.youtube.com/watch?v=yGTUuEx3GkA}{this} webseries
translates really well into multimatrices.

Let's represent the word embeddings of some sentence as $V$ where $\shape{V} = \left<d, n\right>$ where $d$ is the dimensionality of
the embeddings and $n$ is the number of words. Then,

\begin{align*}
  W[i,j]
  &=
  \sum_k
  V[k, i]
  V[k, j] \\
  &=
  \sum_k
  \transpose{V}{1}[i, k]
  V[k, j] \\
  &= (\transpose{V}{1} \mmult{1} V)[i,j]
\end{align*}

And
\begin{align*}
  W'[i,j] &= \frac{
    W[i,j]
  }{
    \sum_k W[k,j]
  } \\
  &= 
  \frac{
    W[i,j]
  }{
   \sum_k 1^{\left<k\right>}[k] W[k,j]
  } \\
  &= 
  \frac{
    W[i,j]
  }{
    (1^{\left<k\right>} \mmult{1} W)[j]
  } \\
  &= 
  W[i,j]
  (1^{\left<k\right>} \mmult{1} W)^{-1}[j] \\
  &= 
  W \dmult{1} (1^{\left<k\right>} \mmult{1} W)^{-1} [i,j]
\end{align*}

And,
\begin{align*}
  Y[i,*]
  &=
  \sum_j W'[i,j] V[j,*] \\
  &=
  (W \mmult{1} V)[i,*]
\end{align*}

So our resulting formula are,
\[ W = \transpose{V}{1} \mmult{1} V \]
\[ W' = W \dmult{1} (1^{\left<k\right>} \mmult{1} W)^{-1} \]
\[ Y = W' \mmult{1} V \]

Or in one line,

\[ Y = \transpose{V}{1} \mmult{1} V \dmult{1} (1^{\left<k\right>} \mmult{1} \transpose{V}{1} \mmult{1} V)^{-1} \mmult{1} V \]

And in the \href{https://www.youtube.com/watch?v=tIvKXrEDMhk}{next} video we see the introduction
of the $W_k$ key, $W_q$, query, and $W_v$ value matrices such that,

\[ W = \transpose{\left(W_q \mmult{1} V\right)}{1} \mmult{1} \left(W_k \mmult{1} V\right) \]
\[ W' = W \dmult{1} (1^{\left<k\right>} \mmult{1} W)^{-1} \]
\[ Y = W' \mmult{1} \left(W_v \mmult{1} V \right) \]

Multiheaded attention simply uses multiple layers of the attention mechanism along with some way to 
reduce the attention down to a single layer.

\[ \shape{W_k} = \shape{W_q} = \shape{W_v} = \left<h,d,d\right> \]
\[ \shape{W} = \shape{W'} = \left<h,n,n\right> \]
\[ \shape{Y} = \left<d,n,h\right> \]
\[ W = \transpose{\left(W_q \mmult{1} V\right)}{2} \mmult{1} \left(W_k \mmult{1} V\right) \]
\[ W[h,*^{\left<n,n\right>}] = \transpose{\left(W_q[h,*^{\left<d,d\right>}] \mmult{1} V\right)}{1} \mmult{1} \left(W_k[h,*^{\left<d,d\right>}] \mmult{1} V\right) \]
\begin{align*}
  W[\bar{h}, n_1, n_2]
  &=
  \sum_k
  (W_q[\bar{h},*] \mmult{1} V)[k, n_1]
  (W_k[\bar{h},*] \mmult{1} V)[k, n_2] \\
  &=
  \sum_{k}
  (W_q \mmult{1} V)[\bar{h}, k, n_1]
  (W_k \mmult{1} V)[\bar{h}, k, n_2] \\
  &=
  \sum_{k,l}
  (W_q \mmult{1} V)[l, k, n_1]
  (W_k \mmult{1} V)[\bar{h}, k, n_2]
  \delta_{\bar{h},l} \\
  &=
  \sum_{k,l}
  \transpose{(W_q \mmult{1} V)}{2}[n_1, l, k]
  \transpose{(W_k \mmult{1} V)}{1}[k, n_2, \bar{h}]
  \delta_{\bar{h},l} \\
  &=
  \sum_{l}
  \left(
    \transpose{(W_q \mmult{1} V)}{2}
    \mmult{1}
    \transpose{(W_k \mmult{1} V)}{1}
  \right)[n_1, l, n_2, \bar{h}] 
  \delta_{\bar{h},l} \\
  &=
  \sum_{l}
  \transpose{\left(
    \transpose{(W_q \mmult{1} V)}{2}
    \mmult{1}
    \transpose{(W_k \mmult{1} V)}{1}
  \right)}{2}[n_2, \bar{h}, n_1, l] 
  \delta_{l, \bar{h}} \\
  &=
  ?
\end{align*}
Ugh, this seems to require a new operation. Because what is,
\[ C[\bar{a}] = \sum_{\forall \bar{c}} A[\bar{a} \oplus \bar{c}] B[\bar{a} \oplus \bar{c}] \]
It's like a mix between the regular and hadamard products.
\[ C = A \otimes B \]
Maybe?
